---
title: "Analysis7"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis7}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Find a claim made in the media made about tropical cyclones and climate change, and check whether the claim is supported or not by your data. You should pick a claim for which your data has relevant information. Cite the media source.

"In the North Atlantic, Studholme et al suggest that we will see expansion of the tropics and a northward shift in tropical cyclones" (Waterman, 2022).

For example, Subtropical Storm Alpha was "the first tropical cyclone to ever make landfall in Portugal and the eastern-most cyclone formation event ever recorded in the North Atlantic" (2022).

*Reference*
Waterman, S. (2022, February 5). Are Atlantic Hurricanes Moving North? Climate Adaptation Center. https://www.theclimateadaptationcenter.org/2022/02/05/are-atlantic-hurricanes-moving-north/

ChatGPT says: "While there is some evidence to suggest a trend toward northward movement in the North Atlantic, it's also possible that other factors could be influencing the behavior of these storms as well."

```{r setup}
library(hurricaneHelper)
data(hurdat)

hurdat$year <- as.numeric(substr(hurdat$date, 1, 4))

# Subset the dataset to include only North Atlantic tropical cyclones
north_atlantic <- subset(hurdat, hurdat$longitude >= -100 &
                             hurdat$longitude <= -20 &
                             hurdat$latitude >= 0 &
                             hurdat$latitude <= 60)

# Calculate the average latitude of maximum intensity for each year
lat_by_year <- aggregate(north_atlantic$latitude, list(north_atlantic$year), max)

# Rename the columns in the aggregate output
names(lat_by_year) <- c("Year", "Avg_Latitude")

# Fit a linear regression model to the average latitude of maximum intensity
# and year, where year is measured in years since 1980
model <- lm(Avg_Latitude ~ Year, data = lat_by_year)

# View the summary of the model
summary(model)

# The estimate found by the model is 0.042412. This means that, on average, the
# latitude of maximum intensity has shifted northward by an average of 0.042412 
# degrees per year since 1851. The p-value is 3.695e-06. Thus, if we use a 
# significance level of 0.05, this means that the relationship if significantly
# significant, providing evidence that the maximum intensity of hurricanes is 
# shifting northward.
```
